---
title: "Swiss Films"
author: "@angelozehr"
date: "July 2019"
---

# 2019-07-reading-pdfs-in-r

Read more about this project in this blogpost:
<https://blog.az.sg/posts/reading-pdfs-in-r>


```{r, echo = FALSE}
package_date <- "2019-03-01" # date of the CRAN snapshot that
R_version <- "3.5.3" # R-Version to use
options(Ncpus = 4) # use 4 cores for parallelized installation of packages
if (R_version != paste0(version$major, ".", version$minor)){
  stop("ERROR: specified R version does not match currently used.")
}
```

This repo is a simplified version of [rddj-template](https://github.com/grssnbchr/rddj-template) by Timo Grossenbacher.

This report was generated on `r Sys.time()`. R version: `r paste0(version$major, ".", version$minor)` on `r version$platform`. For this report, CRAN packages as of `r package_date` were used.


### Original data source

The PDFs in the input folder belong to [Filmdistribution Schweiz](http://filmdistribution.ch/).

## Preparations

```{r, echo = FALSE}
detach_all_packages <- function() {
  basic_packages_blank <-  c("stats",
                             "graphics",
                             "grDevices",
                             "utils",
                             "datasets",
                             "methods",
                             "base")
  basic_packages <- paste("package:", basic_packages_blank, sep = "")

  package_list <- search()[
    ifelse(unlist(gregexpr("package:", search())) == 1, TRUE, FALSE)]

  package_list <- setdiff(package_list, basic_packages)

  if (length(package_list) > 0)  for (package in package_list) {
    detach(package, character.only = TRUE, unload = TRUE)
    print(paste("package ", package, " detached", sep = ""))
  }
}

detach_all_packages()

# this allows multiple persons to use the same RMarkdown
# without adjusting the working directory by themselves all the time
source("scripts/csf.R")
path_to_wd <- csf() # if this - for some reason - does not work,
# replace with a hardcoded path, like so: "~/projects/rddj-template/analysis/"
if ( is.null(path_to_wd) | !dir.exists(path_to_wd)) {
  print("WARNING: No working directory specified for current user")
} else {
  setwd(path_to_wd)
}

# suppress scientific notation
options(scipen = 999)

# unload global rstudioapi and knitr again to avoid conflicts with checkpoint
# this is only necessary if executed within RStudio
# outside of RStudio, namely in the knit.sh script, this causes RMarkdown
# rendering to fail, thus should not be executed there
if (Sys.getenv("RSTUDIO") == "1"){
  detach_all_packages()
}
```


### Packages definieren

```{r, echo=TRUE, message=FALSE, warning=FALSE}
# from https://mran.revolutionanalytics.com/web/packages/checkpoint/vignettes/using-checkpoint-with-knitr.html
# if you don't need a package, remove it from here (commenting is probably not sufficient)
# tidyverse: see https://blog.rstudio.org/2016/09/15/tidyverse-1-0-0/
cat("
library(rstudioapi)
library(tidyverse) # ggplot2, dplyr, tidyr, readr, purrr, tibble
library(glue) # get the tidyverse glue command without namespacing
library(magrittr) # pipes
library(curl) # for downloading pdfs
library(tabulizer) # extract data from pdfs via tabula (requires JAVA)
library(pdftools) # read pdf text without JAVA!
library(tesseract) # OCR for reading non-standard pdfs
library(scales) # scales for ggplot2
library(jsonlite) # json
library(lintr) # code linting
library(sf) # spatial data handling
library(rmarkdown)
library(fuzzyjoin)",
file = "manifest.R")
```

If you encounter problems installing tabulizer (maybe you get an error message like `rJava.so Reason: image not found`), check if you have java installed on your computer by typing `java -version` in your terminal. If nothing gets returned, install Java by following [these instructions](https://blog.az.sg/posts/reading-pdfs-in-r) (at least if you have Mac OS).


### Packages installieren

```{r, echo=TRUE, message=FALSE, warning=FALSE}
# if checkpoint is not yet installed, install it (for people using this
# system for the first time)
if (!require(checkpoint)) {
  if (!require(devtools)) {
    install.packages("devtools", repos = "http://cran.us.r-project.org")
    require(devtools)
  }
  devtools::install_github("RevolutionAnalytics/checkpoint",
                           ref = "v0.3.2", # could be adapted later,
                           # as of now (beginning of July 2017
                           # this is the current release on CRAN)
                           repos = "http://cran.us.r-project.org")
  require(checkpoint)
}
# nolint start
if (!dir.exists("~/.checkpoint")) {
  dir.create("~/.checkpoint")
}
# nolint end
# install packages for the specified CRAN snapshot date
checkpoint(snapshotDate = package_date,
           project = path_to_wd,
           verbose = T,
           scanForPackages = T,
           use.knitr = F,
           R.version = R_version)
rm(package_date)
```


### Packages laden

```{r, echo=TRUE, message=FALSE, warning=FALSE}
source("manifest.R")
unlink("manifest.R")
sessionInfo()
# clean up
rm(path_to_wd, project_name, R_version, user_name, csf, detach_all_packages)
```


## Read Tables

By manually inspecting the files we recognise that the pdfs changed in format once during the time. Namely between week 44 and 45 in year 2010. The content itself did not change, but the column specifications did.

```{r define format_changed}

# define date that differenciates formats in before and after
format_changed <- as.Date("2010-11-01")

pdf_files_to_read <- list.files("input/pdf")

```


```{r first try with tabulizer}

extracted_data <- pdf_files_to_read %>%
  map_df(function(current_filename) {
    current_path <- glue("input/pdf/{current_filename}")

    date_search <- "\\d{2}.\\d{2}.\\d{4}.*\\d{2}.\\d{2}.\\d{4}"
    numbers_search <- "[\\d']+\\s+[\\d']+\\s+[\\d']+$"
    region_search <- "(Deutschschweiz|Suisse Romande|Svizzera Italiana)"

    # try to extract all data into it's own columns
    # by try-catching the error and running the browser
    # we try to find out which files cause an error
    tryCatch({
      # read pdf file as text
      file_content <- extract_text(current_path)

      # if not successfull, return an error entry
      if (file_content == "\n") {
        error_entry <- tibble(
          filename = current_filename,
          error = "File was empty"
        )
        return(error_entry)
      }

      result <- file_content %>%
        str_split("\n") %>%
        as_tibble(.name_repair = make.names) %>%
        mutate(
          dates = str_extract(X, date_search),
          # remove all that is not a number, a dash or a dot
          dates = str_replace_all(dates, "[^\\s\\d\\.]", ""),
          numbers = str_extract(X, numbers_search),
          region = str_extract(X, region_search),
          # remove numbers from source
          X = str_replace(X, numbers_search, ""),
          # remove thousand separator from numbers
          numbers = str_replace_all(numbers, "'", "")
        ) %>%
        # fill up region downwards
        fill(region, dates)

      result %<>%
        # separate dates into from and to
        separate(
          dates,
          sep = "\\s+",
          into = c("from", "to")
        ) %>%
        # convert columns from and to to Dates
        mutate_at(vars(from, to), as.Date, format = "%d.%m.%Y") %>%
        # filter out unused lines
        filter(!is.na(numbers))

      # somehow in separate we cannot work with column names (from)
      # so we need to cancel the pipe here and start a new one
      current_from <- result %>% pull(from) %>% min(na.rm = TRUE)
      current_into <- ifelse(
        current_from < format_changed,
        # what's even more annoying: we cannot save a simple character vector
        # e.g. c("adm", "scr", "tot") here, only the first value would be stored
        "admissions,screens,total",
        "screens,admissions,total"
      )

      result %<>%
        # separate numbers into the three values
        separate(
          numbers,
          sep = "\\s+",
          # the order of the columns is not the same in the old and new columns
          into = str_split(current_into, pattern = ",")[[1]]
        )

      # weird enough, in the old format, the rank is appended to the admissions
      if (current_from < format_changed) {
        result %<>% mutate(
          rank = str_extract(admissions, "([1-9]|10)$"),
          # remove it from the source
          admissions = str_replace(admissions, "([1-9]|10)$", "")
        )
      } else {
        # in the new format, its prepended to the first column
        result %<>% mutate(
          rank = str_extract(X, "^\\d+ "),
          # remove it from the source
          X = str_replace(X, "^\\d+ ", "")
        )
      }

      # return
      result %>%
        rename(movie_and_distributor = X) %>%
        # convert numbers to numeric
        mutate_at(vars(screens, admissions, total, rank), as.numeric) %>%
        # add column for source
        mutate(filename = current_filename)

      },
      error = function(cond) {
        browser()
      }
    )

  })

```

After I tried the `tabulizer` approach I had to realize that it works, but not or all pdfs. Plus we were not able to keep the movie name and the distributor apart. I switched to PDFTools instead.


### PDFTools

The result of `pdf_data` looks not very helpful:

```{r}
knitr::kable(
  pdf_data("input/pdf/top10adm-30_10.pdf")[[1]] %>%
    top_n(10)
)
```

Whereas the result of `pdf_text` is a text file where all the position of a text gets re-drawn as plain text with white spaces. When we inspect a file we see that the numbers of each column align perfectly:

```{r}
pdf_text("input/pdf/top10adm-30_10.pdf") %>%
  str_split("\n") %>%
  print()
```

```{r pdftools}

# prepare empty table for totals per filename
totals_for_quality_check <- tibble()

extracted_data <- pdf_files_to_read %>%
  map_df(function(current_filename) {
    current_path <- glue("input/pdf/{current_filename}")
    current_filecontent <- pdf_text(current_path)

    # save the information about start and end date of observation
    time_info <- current_filecontent %>%
      str_match("(\\d{2}\\.\\d{2}\\.\\d{4}).*?(\\d{2}\\.\\d{2}\\.\\d{4})")

    # if time info is empty, return early, this file is corrupt
    if (is.na(time_info[1, 1])) {
      return(tibble(
        filename = current_filename,
        error = "file corrupt"
      ))
    }

    # extract start and end of observation period (Thu - Sun)
    from <- as.Date(time_info[1, 2], format = "%d.%m.%Y")
    to <- as.Date(time_info[1, 3], format = "%d.%m.%Y")

    # by creating a named character vector we can define the regular expression
    # for the search in a more human readable format
    movie_data_search <- c(
      "region" = "^\\s*(Deutschschweiz|Suisse Romande|Svizzera Italiana)|",
      "rank" = "^\\s*(\\d+)\\s+",
      "title" = "(.*?)\\s+",
      "distributor" = "([\\w]+!?)\\s+",
      "screens" = "(\\d+)\\s+",
      "admissions" = "([\\d']+)\\s+",
      "total" = "([\\d']+)\\s*$"
    )

    current_lines <- current_filecontent %>%
      str_split("\n") %>%
      first() %>%
      as_tibble()

    # save away totals for later quality check
    totals_for_quality_check <<- bind_rows(
      totals_for_quality_check,
      current_lines %>%
        filter(str_detect(value, "^Total Top 10")) %>%
        mutate(filename = current_filename)
    )

    # we match the lines against the search above
    result <- tidyr::extract(
      current_lines,
      1, # from first column
      into = names(movie_data_search),
      # make the regex case insensitive
      regex = paste0("(?i)", glue_collapse(movie_data_search)),
      ignore_case = TRUE
    )

    # return
    result %>%
      fill(region) %>%
      # convert numbers to numeric
      mutate_at(
        vars(rank, screens:total),
        ~ as.numeric(str_replace_all(., "[^\\d]", ""))
      ) %>%
      # filter out unnecessary lines
      filter(!is.na(screens) & !is.na(rank)) %>%
      # add info about week
      mutate(
        from = from,
        to = to,
        filename = current_filename
      )
  })

```


##### Quality check

We saved away all the lines that contain the total admissions per language region. Now we can compare these totals with the sum of the numbers that we extracted above.

```{r}

knitr::kable(
  totals_for_quality_check %>%
    tidyr::extract(
      value,
      into = c("region", "total_is"),
      regex = regex(
        "(?i)(Deutschschweiz|Suisse Romande|Svizzera Italiana).*?([\\d']+)\\s*$",
        ignore_case = TRUE
      )
    ) %>%
    # convert to numeric
    mutate(
      total_is = as.numeric(str_replace_all(total_is, "[^\\d]", ""))
    ) %>%
    left_join(
      extracted_data %>%
        group_by(filename, region) %>%
        summarise(total_should = sum(admissions, na.rm = TRUE)),
      by = c("region", "filename")
    ) %>%
    # show entries where calculated and extracted total do not match
    filter(total_is != total_should)
)

```

The table above should be empty.


#### Unreadable files

But, we see that the following files could not be read:

```{r}

weird_files <- extracted_data %>%
  filter(!is.na(error)) %>%
  pull(filename)

knitr::kable(
  weird_files
)

# throw away these entries and columns as we won't need them anymore
extracted_data %<>%
  filter(is.na(error)) %>%
  select(-error)

```

We read them in via tesseract. This is an OCR engine that converts the file to png first and then reads all the characters in it.


### Tesseract

```{r}

# set up empty list of lines to debug
debug_lines <- tibble()

# the regions cannot be read in because they're on a dark background
# we'll add them differently (see chunks below)
movie_data_search <- c(
  "rank" = "^\\s*(\\d+)\\s+",
  "title" = "(.*?)\\s+",
  "distributor" = "([\\w]+!?)\\s+",
  "screens" = "(\\d+)\\s+",
  "admissions" = "([\\d']+)\\s+",
  "total" = "([\\d'\\.]+)\\s*$"
)

# what follows here is pretty much the same procedure as in pdftools above
# but with some slight differences, so we won't abstract it into a function
extracted_weird_data <- weird_files %>%
  map_df(function(current_filename) {
    current_path <- glue("input/pdf/{current_filename}")

    # if has already been converted to png, take that, otherwise use pdf
    current_path <- ifelse(
      file.exists(str_replace(current_filename, "\\.pdf$", "_1.png")),
      str_replace(current_filename, "\\.pdf$", "_1.png"),
      current_path
    )

    # set up ocr by tesseract: language English
    eng <- tesseract("eng")
    current_filecontent <- ocr(current_path, engine = eng)

    # save the information about start and end date of observation
    time_info <- current_filecontent %>%
      str_match("(\\d{2}\\.\\d{2}\\.\\d{4}).*?(\\d{2}\\.\\d{2}\\.\\d{4})")

    # if time info is empty, return early, this file is corrupt
    if (current_filename == "top10adm-44_06.pdf") {
      # this is the file without a header, so let's just set the dates manually
      from <- as.Date("2006-11-02")
      to <- as.Date("2006-11-05")
    } else if (is.na(time_info[1, 1])) {
      return(tibble(
        filename = current_filename,
        error = "file corrupt"
      ))
    } else {
      # extract start and end of observation period (Thu - Sun)
      from <- as.Date(time_info[1, 2], format = "%d.%m.%Y")
      to <- as.Date(time_info[1, 3], format = "%d.%m.%Y")
    }

    current_lines <- current_filecontent %>%
      str_split("\n") %>%
      first() %>%
      as_tibble()

    # we match the lines against the search above
    result <- tidyr::extract(
      current_lines,
      1, # from first column
      into = names(movie_data_search),
      regex = paste0("(?i)", glue_collapse(movie_data_search))
    )

    # append those that could not be matched to debug variable
    # but add information about filename and from and to as well
    debug_lines <<- bind_rows(
      debug_lines,
      current_lines[is.na(result$rank), ] %>%
      # add info about week
      mutate(
        from = from,
        to = to,
        filename = current_filename
      )
    )

    # return
    result %<>%
      # filter out empty rows
      filter(!is.na(title)) %>%
      # convert numbers to numeric
      mutate_at(
        vars(rank, screens:total),
        ~ as.numeric(str_replace_all(., "[^\\d]", ""))
      ) %>%
      # add region, order is always Deutschschweiz, Romandie, Svizzera Italiana
      mutate(region = case_when(
        rank == 1 & row_number() < 5 ~ "Deutschschweiz",
        rank == 1 & row_number() < 15 ~ "Suisse Romande",
        rank == 1 & row_number() < 25 ~ "Svizzera Italiana"
      )) %>%
      fill(region) %>%
      # add info about week
      mutate(
        from = from,
        to = to,
        filename = current_filename
      )

    # trigger warning if had unsuccessfull lines
    if (nrow(result) != 30) {
      warning(glue(
        "length mismatch in {file}, did not extract 30 movies but {n}\n",
        file = current_filename,
        n = nrow(result)
      ))
    }
    # return
    result
  })

```


#### Obvious Misreads

Unfortunately we get quite some lines with movie data that could not be read:

```{r inspect lines that could not be read}

repair_these_manually <- debug_lines %>%
  # filter out lines that start with Total / Top 10
  filter(!str_detect(value, "^(Total )?(Top 10)")) %>%
  # filter out lines about database and week info and reel
  filter(!str_detect(value, "^(database|Week |reel © )")) %>%
  # filter out empty lines
  filter(nzchar(str_trim(value)))

knitr::kable(
  repair_these_manually
)

write_csv(
  repair_these_manually,
  "output/repair_these_manually.csv"
)

# clean up
rm(weird_files, repair_these_manually)

```

We export them to a file, correct the mistakes by hand and read them in again with the above system. The language region we had to add manually.

```{r}

manually_repaired <- read_csv("input/manually_repaired.csv") %>%
  tidyr::extract(
    value,
    into = names(movie_data_search),
    regex = paste0("(?i)", glue_collapse(movie_data_search))
  ) %>%
  # filter out empty rows
  filter(!is.na(title)) %>%
  # convert numbers to numeric
  mutate_at(
    vars(rank, screens:total),
    ~ as.numeric(str_replace_all(., "[^\\d]", ""))
  )

# append to extracted data frame, but only once!
if (extracted_weird_data %>%
  filter(filename == "top10adm-02_11.pdf") %>%
  nrow() == 29) {
  extracted_weird_data %<>% bind_rows(
    manually_repaired
  )
}

# check if data is complete now
knitr::kable(
  extracted_weird_data %>%
    group_by(filename) %>%
    tally() %>%
    filter(n != 30)
)

# clean up
rm(movie_data_search)

```

The table above should be empty.


#### Quality check

I manually checked the differences in our extracted data and the pdf files and created the file `manually_corrected.csv`. It contains the correct numbers. We'll add them and then do the quality check.

```{r}

extracted_weird_data %<>%
  left_join(
    read_csv("input/manually_corrected.csv"),
    by = c("rank", "region", "filename")
  ) %>%
  # if present, replace with corrected admission
  mutate(
    admissions = if_else(
      is.na(real_admissions),
      admissions,
      real_admissions
    )
  ) %>%
  select(-real_admissions)

# perform again a quality check by using the lines in the debug data frame
debug_lines %>%
  filter(str_detect(value, "^(Total Top 10)")) %>%
  tidyr::extract(
    value,
    into = c("region", "total_is"),
    regex = regex(
      "(?i)(Deutschschweiz|Suisse Romande|Svizzera Italiana).*?([\\d']+)\\s*$",
      ignore_case = TRUE
    )
  ) %>%
  # convert to numeric
  mutate(
    total_is = as.numeric(str_replace_all(total_is, "[^\\d]", ""))
  ) %>%
  left_join(
    extracted_weird_data %>%
      group_by(filename, region) %>%
      summarise(total_should = sum(admissions, na.rm = TRUE)),
    by = c("region", "filename"),
    ignore
  ) %>%
  # show entries where calculated and extracted total do not match
  filter(total_is != total_should)

```

The above table should be empty.

This means we have finally read in all available data and can be pretty sure that all numbers are correct as the sum of each entry and the total in the files match up.


## Linting

This code was linted by [lintr](https://github.com/jimhester/lintr) after he [tidyverse style guide](http://style.tidyverse.org/).

```{r echo=TRUE, message=FALSE, warning=FALSE}
lintr::lint("main.Rmd", linters = lintr::with_defaults())
```
